---
title: "Lab 6"
author: "C.Swick"
date: "2022-11-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(janitor)
library(tidyverse)
library(ggpubr) # for some graphics applications
library(car) # for some ANOVA and regression tests 
```


# Part 2 : Plots and ANOVA assumptions 
```{r, message = FALSE}
penguins <- read_csv(here("data", "penguins.csv"))
```


## Make a subset of penguins data
```{r}
penguins_male <- penguins %>% 
  filter(sex == "male") %>% # filtering out observations
  select(species, body_mass_g) # selecting columns or variable 
```


# qq Plots 

```{r}
# qqplots for each species of penguin (data)

adelie <- penguins_male %>% 
  filter(species == "Adelie")

gentoo <- penguins_male %>% 
  filter(species == "Gentoo")

chinstrap <- penguins_male %>% 
  filter(species == "Chinstrap")


```


```{r}
qq_adelie <- ggplot(adelie, aes(sample = body_mass_g)) + coord_cartesian(ylim = c(3000, 6500)) + 
  geom_qq() + stat_qq() +stat_qq_line() + theme_bw() + labs(title = "Adelie") 

qq_adelie # normal

qq_gentoo <- ggplot(gentoo, aes(sample = body_mass_g))+ coord_cartesian(ylim = c(3000, 6500)) + 
  geom_qq() + stat_qq() +stat_qq_line() + theme_bw() + labs(title = "Gentoo") 

qq_gentoo # normal

qq_chinstrap <- ggplot(chinstrap, aes(sample = body_mass_g))+ coord_cartesian(ylim = c(3000, 6500)) + 
  geom_qq() + stat_qq() +stat_qq_line() + theme_bw() + labs(title = "Chinstrap") 

qq_chinstrap # normal
```


## combine figures 

```{r}
# as a row of figures 

ggarrange(qq_adelie, qq_chinstrap, qq_gentoo, 
          labels = c("A", "B", "C"), ncol = 3, nrow = 1)
```


```{r}
# as a column of figures 

ggarrange(qq_adelie, qq_chinstrap, qq_gentoo, 
          labels = c("A", "B", "C"), ncol = 1, nrow = 3)
```


## Statistical test for normality 

```{r}
shapiro.test(adelie$body_mass_g) # we fail to reject the null, accept that its normally distributed 

shapiro.test(chinstrap$body_mass_g)

shapiro.test(gentoo$body_mass_g)
```


## test of equal variances 

```{r}
# levene's test 

leveneTest(body_mass_g~as.factor(species), data = penguins_male) # fail to reject null, null is that variances are equall 
```


# Part 3A ANOVA 

```{r}
m_penguin_species_results <- aov(body_mass_g~as.factor(species), data = penguins_male)

summary(m_penguin_species_results) # reject the null in favor of alternative, alternative is that at least one mean in species of penguins body mass that is different from the rest 
```



# Part 3b: Multiple Comparisons 

```{r}
TukeyHSD(m_penguin_species_results) # chinstrap-adelie confidence interval passes through 0 so difference is not significant, p value about .05 also shows no significant difference. fail to reject the null that they are different 
```


# Part 3c: plotting ANOVA for publication 

```{r}
# mean plots
ggline(penguins_male, x = "species", y = "body_mass_g", main = "Means and 95% CI for three species of penguins", add=c("mean_ci"), color = "dark grey", ylab = "Body Mass (g)", xlab = "Species", point.color = "red")
```





# Part 4: non parametric tests 
good for skewed data 
violation of normality 
very different variances 
ANOVA should be used over non parametric tests when available because they are more sensitive to true differences 
```{r}
kruskal.test(body_mass_g~as.factor(species), data = penguins_male)
```

## pairwise comparisons using non parametric approach 
```{r}
pairwise.wilcox.test(penguins_male$body_mass_g, penguins_male$species, p.adjust.method = "bonferroni") # p values show differences between adelie and gentoo and chinstrap and gentoo, but no difference between adelie and chinstrap 
# bonferroni adjustment reduces chance of type 1 error but adjusting p value by number of groups being compared 
```

## plot with boxplots 
```{r}
ggboxplot(penguins_male, x = "species", y = "body_mass_g") #thick line shows median of data 
```


## Linear Regression 
```{r, message=FALSE}
trees_mod <- read_csv(here("data", "trees_mod.csv"))

trees <- trees_mod %>%  clean_names()

head(trees)
```



```{r}
tree_scatter <- ggplot(trees, aes(x=height, y=girth))+ geom_point()+ 
  labs(x = "Height (ft)", y = "DBH (in)") +
  xlim(60,90) +ylim(0,25) + theme_bw()

tree_scatter
```


## fit a linear moel 

```{r}
model_1 <- lm(girth~height, data = trees)
model_1 # DBH = -6.2 +.2(height)
summary(model_1)
```

## plot the linear model
```{r}
tree_scatter <- ggplot(trees, aes(x=height, y=girth))+ geom_point()+ 
  labs(x = "Height (ft)", y = "DBH (in)") +
  xlim(60,90) +ylim(0,25) +geom_smooth(method = "lm")+
  theme_bw()

tree_scatter
```






